---
title: "Damselfly Allometry"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

All data prep steps happen in the DataSteps.R file. If you have any questions
it's pretty well annotated.

```{r message = FALSE, warning = FALSE}
source('DataSteps.R')
```
There are a few different ways you could analyze the data. Here's what I've tried
so far. 

## Option 1: Phylogenetic conservatism analysis
Initially it seemed like a good idea to control for relatedness and/or to look
at whether more closely related species had similar growth responses. Unfortunately,
we don't have enough species to pull that off (either within a region or shared across 
regions).

## Option 2: Analyze all data in 1 model
One way to go about this is to include all the lakes in one model and to include
region as a covariate. The downside to this approach is only 3 of the 10 species 
with > 50 observations were found across the midwest and in New England.

```{r}
table(df$Species, df$Region)
```

## Option 3

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
