---
title: "Damselfly Allometry"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

All data prep steps happen in the DataSteps.R file. If you have any questions
it's pretty well annotated.

I dropped species if we did not capture are least 10 individuals at more than 
2 sites. 

```{r message = FALSE, warning = FALSE}
source('DataSteps.R')
```
There are a few different ways you could analyze the data. Here's what I've tried
so far. 

## Option 1: Phylogenetic conservatism analysis
Initially it seemed like a good idea to control for relatedness and/or to look
at whether more closely related species had similar growth responses. Unfortunately,
we don't have enough species to pull that off (either within a region or shared across 
regions).

## Option 2: Analyze all data in 1 model
One way to go about this is to include all the lakes in one model and to include
region as a covariate. The downside to this approach is only 3 of the 7 species 
with that meant minimum observation threshold were found across the midwest and
in Northeast.

```{r echo = FALSE, results = 'asis'}
library('knitr')
temp <- table(df$Species, df$Region)
kable(temp)
```

## Option 3: Analyze each region separately
Another way to approach the problem is to model each region (south central, north central,
and northeast) separately. The advantage here is that we can include all species.
This approach focusses more on growth pattern differnces among species and less
on differences between regions. 


### South Central
```{r echo = FALSE, results = 'asis'}
library('knitr')
temp <-table(df.S$FM_Name, df.S$Species)
kable(temp)
```

### North Central
```{r echo = FALSE, results = 'asis'}
temp <- table(df.N$FM_Name, df.N$Species)
kable(temp)
```

### Northeast
```{r echo = FALSE, results = 'asis'}
library('knitr')
temp <- table(df.E$FM_Name, df.E$Species)
kable(temp)
```

### Step 1: Look at relationship between head width (HW) and outer wing pad length (OWPL)
If correlation perfect = isometry, weak = allometry

I assumed a power law relationship (y = ax^b) based on non-odonate literature.
Under a power law, the exponential (b) is estimated by the slope term. 
But that might not be supported here. Something to think about.

Model: log(HW) ~ Species * log(OWPL) + Sampling Round + (OWPL|Lake)

Looking at slope terms, OWPL and HW allometrically scaled and the strength 
of covariance varied between species, regions, and sampling rounds. 

```{r echo = FALSE, results = 'asis'}
lme.E <- lmer(log(HW) ~  Species*log(OWPL) + as.factor(Sampling.Round) + (log(OWPL)|FM_Name),
             data=df.E)

lme.N <- lmer(log(HW) ~  Species*log(OWPL) + as.factor(Sampling.Round) + (log(OWPL)|FM_Name),
             data=df.N)

lme.S <- lmer(log(HW) ~  Species*log(OWPL) + as.factor(Sampling.Round) + (log(OWPL)|FM_Name),
             data=df.S)

stargazer(lme.S, lme.N, lme.E, 
          single.row = TRUE,
          omit.stat=c("LL","ser","f"), no.space=TRUE,
          intercept.top = TRUE,
          intercept.bottom = FALSE,
          column.labels = c('South Central', 'North Central', 'Northeast'),
          model.numbers = FALSE,
          type = 'html')

```



```{r echo = FALSE}
library(ggplot2)
df.S$Color<-ifelse(df.S$Species == "ENBA", "#e41a1c",
       ifelse( df.S$Species == 'ENEX', '#377eb8',
               ifelse(df.S$Species == 'ENSI', '#4daf4a',
                      ifelse(df.S$Species == 'ENTR', '#984ea3',
                             ifelse(df.S$Species == 'ENVE', '#ff7f00',
                                    ifelse(df.S$Species == 'ENGE', '#ffff33',
                                           ifelse(df.S$Species == 'ENEB_ENHA', '#a65628','#f781bf')))))))

df.N$Color<-ifelse(df.N$Species == "ENBA", "#e41a1c",
                   ifelse( df.N$Species == 'ENEX', '#377eb8',
                           ifelse(df.N$Species == 'ENSI', '#4daf4a',
                                  ifelse(df.N$Species == 'ENTR', '#984ea3',
                                         ifelse(df.N$Species == 'ENVE', '#ff7f00',
                                                ifelse(df.N$Species == 'ENGE', '#ffff33',
                                                       ifelse(df.N$Species == 'ENEB_ENHA', '#a65628','#f781bf')))))))

df.E$Color<-ifelse(df.E$Species == "ENBA", "#e41a1c",
                   ifelse( df.E$Species == 'ENEX', '#377eb8',
                           ifelse(df.E$Species == 'ENSI', '#4daf4a',
                                  ifelse(df.E$Species == 'ENTR', '#984ea3',
                                         ifelse(df.E$Species == 'ENVE', '#ff7f00',
                                                ifelse(df.E$Species == 'ENGE', '#80b1d3',
                                                       ifelse(df.E$Species == 'ENEB_ENHA', '#a65628','#f781bf')))))))


ggplot(data=df.S, aes(y=HW, x=OWPL, colour = Species))+
  # geom_jitter()+
  geom_smooth(method=lm, se=F, lwd=1)+
  scale_x_continuous(trans='log', breaks = c(0.1,0.5,1,2,4,6))+
  scale_y_continuous(trans='log', breaks = c(3,4,6,8))+
  scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  ggtitle("South Central") +
  facet_grid(~Sampling.Round)
#  guides(color=FALSE)


ggplot(data=df.N, aes(y=HW, x=OWPL, colour=Species))+
  # geom_jitter()+
  geom_smooth(method=lm, se=F, lwd=1)+
  scale_x_continuous(trans='log', breaks = c(0.1,0.5,1,2,4,6))+
  scale_y_continuous(trans='log', breaks = c(2,4,6,8))+
  scale_color_manual(values = c('#377eb8', '#4daf4a', '#ff7f00', 'black'))+
  facet_grid(~Sampling.Round)+
  ggtitle("North Central")
#  guides(color=FALSE)


ggplot(data=df.E, aes(y=HW, x=OWPL, colour=Species, group=Species))+
  geom_smooth(method=lm, se=F, lwd=1)+
  scale_x_continuous(trans='log', breaks = c(0.1,0.5,1,2,3))+
  scale_y_continuous(trans='log', breaks = c(1.5,2,3))+
  scale_color_manual(values = c('#377eb8', '#4daf4a', '#ff7f00', 'black', '#f781bf'))+
  ggtitle("Northeast")+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

```

#### Step 2: Explore data to see if any environmental factors were associated with growth strategy

First I checked to see how to best explain the relationship between HW and OWPL

I tested 4 models:  
r1: HW ~ OWPL  
r2: log(HW) ~ log(OWPL)  
r3: log(HW) ~ log(OWPL^2)  
r4: log(HW) ~ log(OWPL) + log(OWPL^2)  

South Central:
```{r echo = FALSE, warning = FALSE, message = FALSE}
source('DataSteps.R')
r1<-lm(HW ~ OWPL, data=df.S)
r2<-lm(lHW ~ lOWPL, data = df.S)
r3<-lm(lHW ~ log(OWPL^2), data=df.S)
r4<-lm(lHW ~ lOWPL + log(OWPL^2), data=df.S)
AIC(r1,r2,r3,r4)
```
North Central:  
```{r echo = FALSE, warning = FALSE, message = FALSE}
source('DataSteps.R')
r1<-lm(HW ~ OWPL, data=df.N)
r2<-lm(lHW ~ lOWPL, data = df.N)
r3<-lm(lHW ~ log(OWPL^2), data=df.N)
r4<-lm(lHW ~ lOWPL + log(OWPL^2), data=df.N)
AIC(r1,r2,r3,r4)
```

Northeast:  
```{r echo = FALSE, warning = FALSE, message = FALSE}
source('DataSteps.R')
r1<-lm(HW ~ OWPL, data=df.E)
r2<-lm(lHW ~ lOWPL, data = df.E)
r3<-lm(lHW ~ log(OWPL^2), data=df.E)
r4<-lm(lHW ~ lOWPL + log(OWPL^2), data=df.E)
AIC(r1,r2,r3,r4)
```

Across the board log(HW) ~ log(OWPL) accounts for the most variance and is
the most parsimonious.

I then extracted the residuals from each model and used these as the response variable
to identify environmental covariates that were associated with more or less investment
in HW than we'd otherwise expect given OWPL.

The model was: Residual HW ~ Species + Competitor Density + Species x Competitor Density + 
Fish Density + Species x Fish Density + Macrophtye Density + Species x Macrophyte Density
+ Prey Density + Species x Prey Density + H2O.PC1 + Species x H2O.PC1 + Sampling.Round +
(1|Lake)

Note: I didn't include chl-a becuase it was colinear with competitor density

South Central:
```{r echo = FALSE, warning = FALSE, message = FALSE}
source('DataSteps.R' )

#Model effects  
covar.lmeS <- lmer(resid.HW ~ Species*Competitorm2 + Species*Fish.Densitym2 +  
                     Species*Shoot.Countm2 + Species*Prey.CPU + Species * H2O.PC1 +
                     Sampling.Round + (1|FM_Name), data=df.S2)  

#summary(covar.lmeS)$coef  

print('South Central')
kable(Anova(covar.lmeS))

ggplot()+
  geom_smooth(data=df.S2, aes(x = Competitorm2, y = resid.HW), method='lm', se = T) +
  # geom_point(data = df.S2, aes(x= Competitorm2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  #scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Competitor density (m3)')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.S2, aes(x = Competitorm2, y = resid.HW, color=Species), method='lm', se = T) +
  # geom_point(data = df.S2, aes(x= Competitorm2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Competitor density (m3)')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.S2, aes(x = Fish.Densitym2, y = resid.HW), method='lm', se = T) +
  # geom_point(data = df.S2, aes(x= Competitorm2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  #scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Fish density (m3)')+
  facet_grid(~Sampling.Round)

ggplot()+
  geom_smooth(data=df.S2, aes(x = Fish.Densitym2, y = resid.HW, color=Species), method='lm', se = T) +
  # geom_point(data = df.S2, aes(x= Competitorm2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Fish density (m3)')+
  facet_grid(~Sampling.Round)
#  guides(color=FALSE)
```

North Central and northeast models were rank deficient, but if combined data into a North model:
```{r echo = FALSE, warning = FALSE, message = FALSE}
source('DataSteps.R' )
# covar.lmeN <- lmer(resid.HW ~ Species*Competitorm2 + Species*Fish.Densitym2 +  
#                      Species*Shoot.Countm2 + Species*Prey.CPU + Species * H2O.PC1 +
#                      Sampling.Round + (1|FM_Name), data=df.N2)
# summary(covar.lmeN)
# Anova(covar.lmeN)
# 
# #Insufficient ENAS, ENMI, and ENTR to include
# covar.lmeE <- lmer(resid.HW ~ Species*Competitorm2 + Species*Fish.Densitym2 +  
#                      Species*Shoot.Countm2 + Species*Prey.CPU +SamplingRound + (1|FM_Name), data=df.E2)
# summary(covar.lmeE)
# Anova(covar.lmeE)

#Combine both northern samples
#No collinearity problems.
covar.lmeAllN <- lmer(resid.HW ~ Species*Competitorm2 + Species*Fish.Densitym2 +  
                     Species*Shoot.Countm2 + Species*Prey.CPU + Species * H2O.PC1 +
                     Sampling.Round + (1|FM_Name), data=df.NAll2)
#summary(covar.lmeAllN)
#View(summary(covar.lmeAllN)$coef)
kable(Anova(covar.lmeAllN))

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = Shoot.Countm2, y = resid.HW), method='lm', se = T) +
  # geom_point(data = df.S2, aes(x= Shoot.Countm2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  #scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Shoot count (m3)')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = Shoot.Countm2, y = resid.HW, colour = Species), method='lm', se = T) +
  # geom_point(data = df.NAll2, aes(x= Shoot.Countm2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Shoot count (m3)')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = Fish.Densitym2, y = resid.HW), method='lm', se = T) +
  # geom_point(data = df.NAll2, aes(x= Fish.Densitym2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  #scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Fish density (m3)')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = Fish.Densitym2, y = resid.HW, colour = Species), method='lm', se = T) +
  # geom_point(data = df.NAll2, aes(x= Fish.Densitym2, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Fish density (m3)')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = Prey.CPU, y = resid.HW), method='lm', se = T) +
  # geom_point(data = df.NAll2, aes(x= Prey.CPU, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  #scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Prey CPU')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = Prey.CPU, y = resid.HW, colour = Species), method='lm', se = T) +
  # geom_point(data = df.NAll2, aes(x= Prey.CPU, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Prey CPU')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)

ggplot()+
  geom_smooth(data=df.NAll2, aes(x = H2O.PC1, y = resid.HW), method='lm', se = T) +
  # geom_point(data = df.NAll2, aes(x= H2O.PC1, y = resid.HW, color = Species))+
  # coord_cartesian(ylim = c(-0.25,0.25))+
  #scale_color_manual(values = c("#e41a1c", '#80b1d3', '#4daf4a', '#984ea3','#ff7f00'))+
  scale_y_continuous('Residual HW') + 
  scale_x_continuous('Water PC1')+
  facet_grid(~Sampling.Round)
  #guides(color=FALSE)
```
